What's hot and what's not:\\Windowed developer topic analysis

### 2009-04-12 lda-paper.tex
One good edit pass

\shrinkit
\section{Methodology}
\shrinkit
\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
that fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.  The
compact-trend-view is convenient because it shows all of the topic
information, as shown by the zoomed-in view of MaxDB 7.500's
compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
repeating continuous trends easy to pick out, although discontinuous
trends are harder to spot.

@@

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most common trend, which
stretched across the entire development, contained these same words
(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
mentions Dal Zotto, while the third largest trend mentions the
\texttt{[PATCHES]} mailinglist and the names of some patch
contributors.  Other repeating topics refer to portability with Win32,
Central Europe Time (CEST) from email headers, issues with ALTER TABLE
and CVS branch merging, because CVS does not record merges explicitly.

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs dominated unique topics during some months. For
instance in the last month of MaxDB 7.500 development every topic
contained 1 unique Bug tracker URL, this pattern did not occur in the
previous month. We investigated the revisions and we did find
there was a difference in behavior between, and that the last month of
development commits were mostly referencing the bug tracker. If the
topics of one month were about unique bug tickets being
addressed, the global topic analysis would probably miss this, yet these
bug tickets were the focus of development for that month but not
necessarily globally relevant.

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur four
times, yet in the global-trend-view (Figure \ref{fig:statictopics}) it
is not in any of the topics. A query optimizer is rather important to
an DBMS, but as we have shown it just does not pop up as a topic on
its own. We tried to remove words to see if we could get an 
\emph{optimizer} topic. After removing stop words and then two of the most
common words, the global analysis finally had a topic with optimizer in
its topic ten words. This shows that optimizer was important but it
had been obscured, while it would have been noticed using the more
locally relevant approach that we propose.

We noticed that commits that mentioned \emph{PHP} occurred two thirds
less frequently than commits that mentioned \emph{Perl}, but
\emph{Perl}-related topics appeared in the global static topics for
MaxDB while \emph{PHP}-related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different windowed topics and 1 global
topic. Perhaps this is because there was a cluster of \emph{Perl}
mentions during one month while the \emph{PHP} mentions were more
spread out.

We noticed that \emph{Perforce} and \emph{Changelist} were in just
about every topic, so we added them to the stop words list. As a
result, longer trends were shortened and sometimes the total
number of topics found per month were less.  Evaluating different
similarity thresholds showed that by removing common words one reduces
the general similarity of topics. That said, the larger topics were still
clearly apparent. Thus if more relevant stop words are added one
should tune the topic similarity parameters to handle these changes.

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a specified number of topics and then tracked them over
time. This means that a single, static set of topics was tracked
across the whole development history of a project. \emph{Global topic
  analysis} is topic analysis applied against the entire lifetime of a
project.

We carried out the same kind of analysis,          to compare the
results against our time-windowed approach.  To produce
figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
number of messages per month that were related to that topic. We found
was that one topic would often dominate the results, as shown in the
third row of Figure \ref{fig:statictopics}, while messages related to
other topics did not appear often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

We truncated the topics to top 10 tokens, so this summary of a topic
might not have been as useful as determining the actual topic distance
between two word topic distributions. By truncating tokens we could be
losing valuable data.

Each project seemed to have their own set of stop words, and we did not
remove them, but perhaps dropping some of these stop words would aid
the clarity of topics produced through topic analysis; alternatively
it might cause our topic similarity plots to be fundamentally
different.

@@

We presented several visualization techniques that focused on
different aspects of the data: temporality of trends, trend size, and
a compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics, as
well, it shows topics by similarity so one can track trends across
time.  Our trend-histogram immediately highlights and measures the
size of trends while our trend-time-line view shows how a topic
reoccurs over time. These visualizations help us get an general
feeling about the common and unique topics that developers focus
on during development. If implemented properly a user could easily zoom in and query a
summary of a topic or trend.

\shrinkit
\subsection{ Future Work}
\shrinkit

@@

\shrinkit
\bibliographystyle{latex8}
\bibliography{lda-paper}

>>>>

\shrinkit
\section{Methodology}
\shrinkit

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking, in practice
there are lot of small trends (90\% of all trends contain 1 topic)
that fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.  The
compact-trend-view is convenient because it shows all of the topic
information, as shown by the zoomed-in view of MaxDB 7.500's
compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
repeating continuous trends easy to pick out, although discontinuous
trends are harder to spot.

@@

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most common trend, which stretched across
the entire development, contained these same words (\emph{patch},
\emph{fix}, \emph{update}). The second largest trend mentions Dal
Zotto, while the third largest trend mentions the \texttt{[PATCHES]}
mailinglist and the names of some patc contributors.  Other repeating
topics refer to portability with Win32, Central Europe Time (CEST)
from email headers, issues with ALTER TABLE and CVS branch merging
(CVS does not record merges explicitly).

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs dominated unique topics during some months. For
instance in the last month of MaxDB 7.500 development every topic
contained 1 unique Bug tracker URL, this pattern did not occur in the
previous month. We investigated the revisions and we found that
developers were referencing the bug tracker more during the last
month.  If the topics of one month were about unique bug tickets being
addressed, the global topic analysis would probably miss this, yet
these bug tickets were the focus of development for that month but not
necessarily globally relevant.

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. This shows
that optimizer was important but it had been obscured, but would have
been noticed using the more local topic analysis.

We noticed that commits that mentioned \emph{PHP} occurred two thirds
less frequently than commits that mentioned \emph{Perl}, but
\emph{Perl}-related topics appeared in the global static topics for
MaxDB while \emph{PHP}-related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different topics and 1 global
topic. Perhaps this is because there was a cluster of \emph{Perl}
mentions during one month while the \emph{PHP} mentions were more
spread out.

Just about every topic included the words \emph{Perforce} and
\emph{Changelist}, so we added them to the stop words list. As a
result, longer trends were shortened and sometimes the total number of
topics found per month were reduced.  Evaluating different similarity
thresholds showed that by removing common words one reduces the
general similarity of topics. That said, the larger topics were still
clearly apparent. Thus if more relevant stop words are added one
should tune the topic similarity parameters to handle these changes.

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a specified number of topics and then tracked them over
time. This means that a single, static set of topics was tracked
across the whole development history of a project, we refer to this as \emph{global topic
  analysis}.% is topic analysis applied against the entire lifetime of a
%project.

We carried otu \emph{global topic analysis} on MaxDB 7.500 and
compared this against our windowed topic analysis.  To produce
figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
number of messages per month that were related to that topic. We found
was that one topic would often dominate the results, as shown in the
third row of Figure \ref{fig:statictopics}, while messages related to
other topics did not appear often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a windowed topic analysis gives a much stronger
indication of the fleeting importance of such topics, and can help to
put such a topic into its proper context.

@@

We compared topics using the top 10 tokens, this approach could be
throwing data away and might not be as useful as determining the
actual distance between two word topic distributions.

% We truncated the topics to the top 10 tokens, so this summary of a topic
% might not have been as useful as determining the actual topic distance
% between two word topic distributions. By truncating tokens we could be
% losing valuable data.

Adding and removing stop words produced different results. Our choice
of stopwords could be biased, and could affect the results.
%, and our bias for or against such stop words could affect
%the results.

% Each project seemed to have their own set of stop words, and we did not
% remove them, but perhaps dropping some of these stop words would aid
% the clarity of topics produced through topic analysis; alternatively
% it might cause our topic similarity plots to be fundamentally
% different.

@@

We presented several visualization techniques that focused on
different aspects of the data: temporality of trends, trend size, and
a compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics, as
well, it shows topics by similarity so one can track trends across
time.  Our trend-histogram immediately highlights and measures the
size of trends while our trend-time-line view shows how a topic
reoccurs over time. These visualizations help us get a general feeling
about the common and unique topics that developers focus on during
development. If implemented interactively a user could easily zoom in
and query for a summary of a topic or trend.

\shrinkit
\subsection{ Future Work}
\shrinkit
\shrinkit

@@

\shrinkit
\bibliographystyle{latex8}

\shrinkit
\bibliography{lda-paper}
